---
title: "Lily"
output: html_document
date: "2023-04-15"
---

This is Lily's markdown file

```{r}
library(tidyverse)
library(tidycensus)
library(tidyr)
library(usmap)
library(ggplot2)
library(dplyr)
```

```{r}
attorneys <- read_csv("../data/DataFest_2023_Data/attorneys.csv")

attorneys_time_entries <- read_csv("../data/DataFest_2023_Data/attorneytimeentries.csv")

categories <- read_csv("../data/DataFest_2023_Data/categories.csv")

clients <- read_csv("../data/DataFest_2023_Data/clients.csv")

question_posts <- read_csv("../data/DataFest_2023_Data/questionposts.csv")

questions <- read_csv("../data/DataFest_2023_Data/questions.csv")

state_sites <- read_csv("../data/DataFest_2023_Data/statesites.csv")

subcategories <- read_csv("../data/DataFest_2023_Data/subcategories.csv")

reporting_policy <- read_csv("../data/reporting_policy.csv")
```

I'm interested in the overall demographics of the clients, and also how that correlates with the type of questions that are being asked

The following focuses only on the gender. It is important to note I'm only selecting values with *one* gender value, so I'm filtering out responses from people who put down multiple genders

```{r}
clients %>%
  count(Gender)

combined <- questions %>%
  drop_na(AskedByClientUno) %>%
  right_join(clients, by = c('AskedByClientUno' = 'ClientUno'))

combined

gender_v_category <- combined %>%
  select("Category", "Gender") %>%
  group_by(Gender) %>%
  mutate(
    total_by_gender = n()
  ) %>%
  group_by(Category) %>%
  mutate(
    count = n(),
    prop = count / total_by_gender
  )

gender_v_category

# gender_v_category %>%
#   filter(Gender == c("Female", "Male")) %>%
#   ggplot() +
#     geom_col(aes(x = Category, y = prop, fill = Gender), position = "dodge") +
#     # geom_col(data = male_category, mapping = aes(Category, n), fill = "blue") +
#     coord_flip()

```

I'm going to filter the questions by keywords, because there are a lot of NA categories. The ones i'm interested in are: \_\_\_\_\_\_

```{r}

```

Other topics:

-   Smth ab incorrectly categorized questions - how this relates to proportion of unanswered questions, which categories have more flaws and should be reworded

```{r}
null_prop <- questions %>%
  left_join(clients, by = c("AskedByClientUno" = "ClientUno", "StateAbbr")) %>%
  group_by(StateAbbr) %>%
  mutate(
    is_null = 1 * (TakenByAttorneyUno == "NULL"),
  ) %>%
  count(is_null) %>%
  pivot_wider(
    names_from = "is_null",
    names_prefix = "null",
    values_from = "n"
  ) %>%
  mutate(
    null_percent = null1 / (null0 + null1)
  ) %>%
  arrange(desc(null_percent))

null_prop
```

```{r}
null_prop %>%
  ggplot() +
  geom_col(aes(StateAbbr, null_percent))
```

Try to add counties

```{r}
states_and_fips <- fips_codes %>%
  mutate(
    fips = paste0(state_code, county_code)
  ) %>%
  select(state, county, fips) %>%
  mutate(county = str_remove_all(county, " County"))

states_and_fips
```

```{r}
# test that combining zip and fips works
questions %>%
  left_join(select(clients, -Id), by = c("StateAbbr", "AskedByClientUno" = "ClientUno")) %>%
  group_by(County, StateAbbr) %>%
  left_join(states_and_fips, by = c("County" = "county", "StateAbbr" = "state")) %>%
  filter(County == "Calhoun", StateAbbr == "AL")
```

```{r}
questions_by_county <- questions %>%
  left_join(select(clients, -Id), by = c("StateAbbr", "AskedByClientUno" = "ClientUno")) %>%
  group_by(County, StateAbbr) %>%
  mutate(
    is_null = 1 * (TakenByAttorneyUno == "NULL"),
  ) %>%
  count(is_null) %>%
  pivot_wider(
    names_from = "is_null",
    names_prefix = "null",
    values_from = "n"
  ) %>%
  mutate(
    null_percent = null1 / (null0 + null1)
  ) %>%
  drop_na() %>%
  left_join(states_and_fips, by = c("County" = "county", "StateAbbr" = "state")) %>%
  arrange(desc(null_percent)) %>%
  drop_na()

questions_by_county
```

## Important stuff below

### Entire US county map - nonresponse rate

```{r}
plot_usmap(regions = "counties", data = questions_by_county, values = "null_percent", color = NA) +
  scale_fill_continuous(
    low = "white", high = "red", name = "Nonresponse rate", label = scales::comma
  ) +
  labs(
    title = "Noresponse Rates by US Counties",
    # subtitle = "This is a blank map of the counties of the United States."
  ) +
  theme(legend.position = "right")

```

### Zooming into our 5 case study states (Florida, Texas, Illinois, Indiana, Georgia)

```{r}
plot_usmap(include = c("FL", "TX", "IL", "IN", "GA"), regions = "counties", data = questions_by_county, values = "null_percent", color = NA) +
  scale_fill_continuous(
    low = "white", high = "red", name = "Nonresponse rate", label = scales::comma
  ) +
  labs(title = "Nonresponse rates for 10 case study states") +
  theme(legend.position = "bottom")
```

### Comparing Florida and Texas proportions of non-response rates over time

METHINKS THIS IS A VERY USEFUL PLOT!!!!!!!

Run the cell below to store all the data. theres probably a more efificent way to do this than the 200000 lines of code i have but it is what it is

```{r}
# texas
texas_ignored <- questions %>%
  filter(StateAbbr == "TX", TakenOnUtc == "NULL") %>%
  separate(col = AskedOnUtc, into = c("Date", "ignore1"), sep = -12) %>%
  select(
    Date, TakenOnUtc, -ignore1
  ) %>%
  count(Date)


texas_asked_n <- questions %>%
  separate(col = AskedOnUtc, into = c("AskedOnTX", "ignore2"), sep = -12) %>%
  select(
    StateAbbr,
    AskedOnTX,
    -ignore2
  ) %>%
  filter(StateAbbr == "TX") %>%
  select(
    -StateAbbr
  ) %>%
  count(AskedOnTX) %>%
  rename(
    Date = "AskedOnTX",
    AskedOnTX = "n"
  )


# florida
florida_ignored <- questions %>%
  filter(StateAbbr == "FL", TakenOnUtc == "NULL") %>%
  separate(col = AskedOnUtc, into = c("Date", "ignore1"), sep = -12) %>%
  select(
    Date, TakenOnUtc, -ignore1
  ) %>%
  count(Date)

florida_asked_n <- questions %>%
  separate(col = AskedOnUtc, into = c("AskedOnFL", "ignore2"), sep = -12) %>%
  select(
    StateAbbr,
    AskedOnFL,
    -ignore2
  ) %>%
  filter(StateAbbr == "FL") %>%
  select(
    -StateAbbr
  ) %>%
  count(AskedOnFL) %>%
  rename(
    Date = "AskedOnFL",
    AskedOnFL = "n"
  )


# illinois
illinois_ignored <- questions %>%
  filter(StateAbbr == "IL", TakenOnUtc == "NULL") %>%
  separate(col = AskedOnUtc, into = c("Date", "ignore1"), sep = -12) %>%
  select(
    Date, TakenOnUtc, -ignore1
  ) %>%
  count(Date)

illinois_asked_n <- questions %>%
  separate(col = AskedOnUtc, into = c("AskedOnIL", "ignore2"), sep = -12) %>%
  select(
    StateAbbr,
    AskedOnIL,
    -ignore2
  ) %>%
  filter(StateAbbr == "IL") %>%
  select(
    -StateAbbr
  ) %>%
  count(AskedOnIL) %>%
  rename(
    Date = "AskedOnIL",
    AskedOnIL = "n"
  )

# indiana
indiana_ignored <- questions %>%
  filter(StateAbbr == "IN", TakenOnUtc == "NULL") %>%
  separate(col = AskedOnUtc, into = c("Date", "ignore1"), sep = -12) %>%
  select(
    Date, TakenOnUtc, -ignore1
  ) %>%
  count(Date)

indiana_asked_n <- questions %>%
  separate(col = AskedOnUtc, into = c("AskedOnIN", "ignore2"), sep = -12) %>%
  select(
    StateAbbr,
    AskedOnIN,
    -ignore2
  ) %>%
  filter(StateAbbr == "IN") %>%
  select(
    -StateAbbr
  ) %>%
  count(AskedOnIN) %>%
  rename(
    Date = "AskedOnIN",
    AskedOnIN = "n"
  )

# georgia
georgia_ignored <- questions %>%
  filter(StateAbbr == "GA", TakenOnUtc == "NULL") %>%
  separate(col = AskedOnUtc, into = c("Date", "ignore1"), sep = -12) %>%
  select(
    Date, TakenOnUtc, -ignore1
  ) %>%
  count(Date)

georgia_asked_n <- questions %>%
  separate(col = AskedOnUtc, into = c("AskedOnGA", "ignore2"), sep = -12) %>%
  select(
    StateAbbr,
    AskedOnGA,
    -ignore2
  ) %>%
  filter(StateAbbr == "GA") %>%
  select(
    -StateAbbr
  ) %>%
  count(AskedOnGA) %>%
  rename(
    Date = "AskedOnGA",
    AskedOnGA = "n"
  )



texas_proportions <- texas_asked_n %>%
  left_join(texas_ignored, by = "Date") %>%
  rename(IgnoredTX = "n") %>%
  replace(is.na(.), 0) %>%
  mutate(
    propTX = IgnoredTX / AskedOnTX
  )

florida_proportions <- florida_asked_n %>%
  left_join(florida_ignored, by = "Date") %>%
  rename(IgnoredFL = "n") %>%
  replace(is.na(.), 0) %>%
  mutate(
    propFL = IgnoredFL / AskedOnFL
  )

illinois_proportions <- illinois_asked_n %>%
  left_join(illinois_ignored, by = "Date") %>%
  rename(IgnoredIL = "n") %>%
  replace(is.na(.), 0) %>%
  mutate(
    propIL = IgnoredIL / AskedOnIL
  )

indiana_proportions <- indiana_asked_n %>%
  left_join(indiana_ignored, by = "Date") %>%
  rename(IgnoredIN = "n") %>%
  replace(is.na(.), 0) %>%
  mutate(
    propIN = IgnoredIN / AskedOnIN
  )

georgia_proportions <- georgia_asked_n %>%
  left_join(georgia_ignored, by = "Date") %>%
  rename(IgnoredGA = "n") %>%
  replace(is.na(.), 0) %>%
  mutate(
    propGA = IgnoredGA / AskedOnGA
  )


combined_proportions <-
  texas_proportions %>%
  left_join(florida_proportions, by = "Date") %>%
  left_join(illinois_proportions, by = "Date") %>%
  left_join(indiana_proportions, by = "Date") %>%
  left_join(georgia_proportions, by = "Date") %>%
  select(Date, propTX, propFL, propIL, propIN, propGA) %>%
  rename(TX = "propTX", FL = "propFL", IL = "propIL", IN = "propIN", GA = "propGA") %>%
  pivot_longer(
    cols = c(TX, FL, IL, IN, GA),
    names_to = "state",
    values_to = "proportion"
  ) %>%
  mutate(
    below_threshold = proportion <= 0.1
  )
```

```{r}
combined_proportions %>%
  ggplot() +
  geom_col(aes(x = Date, y = proportion, fill = below_threshold)) +
  facet_wrap(~state, ncol = 2) +
  scale_x_discrete(breaks = c("2018-01", "2019-01", "2020-01", "2021-01")) +
  ggtitle("Proportion of Asked but Unanswered Questions for 5 states")
```

Interesting points:

-   late 2018 was a huge spike for both - June/July-ish

-   both were at a low in beginning of 2020 and have been increasing ever since

    -   check min value for texas (should be in 2020)

Now we will check which states are relevant, meaning they started the program before Jan 2020, and have \>3000 responses in the 2 year time range

```{r}
total <- questions %>%
  separate(col = AskedOnUtc, into = c("Year", "Month", "ignore"), sep = "-", remove = FALSE) %>%
  select(Year, Month, StateAbbr, AskedOnUtc, TakenOnUtc) %>%
  filter(Year == "2020" | Year == "2021" | Year == "2022") %>%
  count(StateAbbr) %>%
  rename(total_amt = "n")

total_null <- questions %>%
  separate(col = AskedOnUtc, into = c("Year", "Month", "ignore"), sep = "-", remove = FALSE) %>%
  select(Year, Month, StateAbbr, AskedOnUtc, TakenOnUtc) %>%
  filter(Year >= "2020", TakenOnUtc == "NULL") %>%
  count(StateAbbr) %>%
  rename(null_amt = "n")
```

```{r}
relevant_states <-
  questions %>% 
  group_by(StateAbbr) %>% 
  summarize(start = min(AskedOnUtc)) %>% 
  filter(start < "2020-01-01 00:00:00") %>%
  pull(StateAbbr)

total %>%
  filter(
    StateAbbr %in% relevant_states,
    total_amt > 3000
  ) %>%
  left_join(total_null, by = "StateAbbr") %>%
  mutate(prop = null_amt / total_amt) %>%
  arrange(desc(prop))

total %>%
  filter(
    StateAbbr %in% relevant_states,
    total_amt > 3000
  ) %>%
  left_join(total_null, by = "StateAbbr") %>%
  mutate(prop = null_amt / total_amt) %>%
  arrange(prop)
```

## #Active attorneys in our 15 states

```{r}

our_15_states <- c("TN", "WI", "FL", "NC", "MO", "VA", "NY", "SC", "IL", "MA", "IN", "TX", "OK", "AZ", "GA")

active_attorneys <- attorneys_data %>%
  filter(
    CreatedUtc > "2020-01-01 00:00:00",
    StateAbbr %in% our_15_states
  ) %>%
  count(StateAbbr)
active_attorneys
```

Using the dataset we curated, make a map plot based on whether pro bono is mandatory in each state

```{r}
df_responses <-
  questions %>% 
  filter(StateAbbr %in% relevant_states) %>%
  group_by(StateAbbr) %>% 
  mutate(
    response_received = 1 * (TakenByAttorneyUno != "NULL")
  ) %>% 
  count(response_received) %>%
  pivot_wider(
    names_from = "response_received",
    names_prefix = "response_",
    values_from = "n"
  ) %>%
  rename(
    responses = response_1,
    nonresponses = response_0
  ) %>% 
  mutate(
    total_questions = nonresponses + responses,
    response_rate = responses / total_questions
  ) %>% 
  left_join(reporting_policy, by = c("StateAbbr" = "State"))

df_responses
```

Now plot using df_responses

```{r}
plot_usmap(include = our_15_states, data = rename(df_responses, state = "StateAbbr"), values = "Reporting") +
  scale_fill_discrete() +
  labs(title = "Pro bono policies for the 15 states") +
  theme(legend.position = "top")
```

Using the dataset we curated, make a map plot based on how many active attorneys there are in each of the 15 states

```{r}
plot_usmap(include = our_15_states, data = rename(active_attorneys, state = "StateAbbr"), values = "n") +
  scale_fill_continuous(
    low = "white", high = "red", name = "#active attorneys"
  ) +
  labs(title = "Number of active attorneys for the 15 relevant states") +
  theme(legend.position = "bottom")
```
